Accurate plant pathogen effector protein classification _ab initio_ with **deepredeff**, an ensemble of convolutional neural networks.

### 2020-06-08 main.Rmd
Move figures inline. Add long tables to supplement

@@

@@

-->
```{r effectorcollection, out.width='50%', fig.align='center', fig.cap="Workflow diagram for collection of effector sequences from the PHI-Base database annotation and cross reference to UniProt"}
knitr::include_graphics(
  "figures/flowchart_effectorsequence_collection.pdf"
)
```

```{r noneffectorcollection, fig.cap="Workflow diagram for collection of secreted non-effector sequences from Ensembl Bacteria, Fungi, and Protists"}
knitr::include_graphics(
  "figures/flowchart_secreted_noneffectorsequence_collection.pdf"
)
```

```{r modelarch, fig.cap="Overview of model architectures tested in this analysis, A: CNN-LSTM model; B: CNN-GRU model; C: LSTM-Embedding model; D: GRU-Embedding model"}
knitr::include_graphics("figures/model_arch.pdf")
```

@@

```{r ensembleresults, fig.cap="Final performance of Ensemble models in each taxon on the unseen, hold-out test data fraction of sequences."}
# apa_table(
#   readr::read_tsv("tables/ensemble_model_performance.tab"),
#   digits = 3,
#   align = 'c',
#   caption = 'Final performance of Ensemble models in each taxon on the unseen, hold-out test data fraction of sequences.'
readr::read_tsv("tables/ensemble_model_performance_tidy.tab") %>%
  ggplot2::ggplot() +
  ggplot2::aes(x = Combination, y = Accuracy) +
  ggplot2::geom_point(ggplot2::aes(colour = EnsembleFunction,
                                   shape = EnsembleFunction), size = 2) +
  ggplot2::scale_shape_manual(values = c("Ensemble Voting" = 19, "Ensemble Weighted" = 17)) +
  ggplot2::scale_color_viridis_d(begin = 0.65, end = 0.1) +
  ggplot2::labs(color = "Ensemble Function", shape = "Ensemble Function") +
  ggplot2::facet_wrap(~ Taxon, scales = "free_x") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

```{r vsothersmetrics, fig.cap="Accuracy and F1 score of the classifications made by our models compared with those made by widely used software", echo=FALSE, message=FALSE}
data.table::fread("data/processed/acc_f1score_vs_othertools.csv") %>%
  dplyr::mutate(model = case_when(
    model == "ensemble_weighted" ~ "Ensemble Weighted",
    model == "cnn_lstm" ~ "CNN-LSTM",
    model == "deepT3" ~ "DeepT3",
    model == "effectiveT3" ~ "EffectiveT3",
    model == "effectorp_1" ~ "EffectorP 1",
    model == "effectorp_2" ~ "EffectorP 2",
    TRUE ~ "EffectR"
  )) %>%
  dplyr::mutate(
    model = factor(model, levels = c("Ensemble Weighted", "DeepT3", "EffectiveT3",
                                    "CNN-LSTM", "EffectorP 1", "EffectorP 2",
                                    "EffectR"))
  ) %>%
  ggplot2::ggplot() +
  ggplot2::aes(
    x = model,
    y = .estimate,
    color = .metric,
    shape = .metric
  ) +
  ggplot2::scale_shape_manual(values = c("Accuracy" = 19, "F1 Score" = 17)) +
  ggplot2::geom_point(size = 2) +
  ggplot2::facet_wrap(~groups, nrow = 1, scales = "free_x") +
  ggplot2::scale_color_viridis_d(begin = 0.65, end = 0.1) +
  ggplot2::labs(x = "Classifier", y = "Score", color = "Metric", shape = "Metric") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

```{r vsotherscfmat, fig.width=9, fig.height=8, fig.cap="Confusion matrices of classifications made by our models compared with those made by widely used software. A: Tools evaluated on bacterial sequences; B: Tools evaluated on fungal sequences; C: Tools evaluated on oomycete sequences.", message = FALSE, echo=FALSE}

@@

```{r activations, fig.height=6, fig.width=10, fig.cap="Grad-CAM activation profiles of CNN-LSTM models trained on when run with all datasets; and sequence logo plots of sequence in the activation region within the dotted lines for effector data and non-effector data. A: Models trained on bacterial sequences; B: Models trained on fungal sequences; C: Models trained on oomycete sequences."}
# knitr::include_graphics("figures/convolution_maps_fixed.pdf")

>>>>

```{r effectorcollection, fig.width=10, fig.height=10, out.width='50%', fig.align='center', fig.cap="Workflow diagram for collection of effector sequences from the PHI-Base database annotation and cross reference to UniProt"}
knitr::include_graphics(
  "figures/flowchart_effectorsequence_collection.pdf"
)
```

```{r noneffectorcollection, fig.width=10, fig.height=10, fig.cap="Workflow diagram for collection of secreted non-effector sequences from Ensembl Bacteria, Fungi, and Protists"}
knitr::include_graphics(
  "figures/flowchart_secreted_noneffectorsequence_collection.pdf"
)
```

@@

```{r modelarch, fig.width=10, fig.height=10, fig.cap="Overview of model architectures tested in this analysis, A: CNN-LSTM model; B: CNN-GRU model; C: LSTM-Embedding model; D: GRU-Embedding model"}
knitr::include_graphics("figures/model_arch.pdf")

@@

@@

```{r ensembleresults, fig.width=8, fig.height=8, fig.cap="Final performance of Ensemble models in each taxon on the unseen, hold-out test data fraction of sequences."}
# apa_table(
#   readr::read_tsv("tables/ensemble_model_performance.tab"),
#   digits = 3,
#   align = 'c',
#   caption = 'Final performance of Ensemble models in each taxon on the unseen, hold-out test data fraction of sequences.'
readr::read_tsv("tables/ensemble_model_performance_tidy.tab") %>%
  ggplot2::ggplot() +
  ggplot2::aes(x = Combination, y = Accuracy) +
  ggplot2::geom_point(ggplot2::aes(colour = EnsembleFunction,
                                   shape = EnsembleFunction), size = 2) +
  ggplot2::scale_shape_manual(values = c("Ensemble Voting" = 19, "Ensemble Weighted" = 17)) +
  ggplot2::scale_color_viridis_d(begin = 0.65, end = 0.1) +
  ggplot2::labs(color = "Ensemble Function", shape = "Ensemble Function") +
  ggplot2::facet_wrap(~ Taxon, scales = "free_x") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

```{r vsothersmetrics, fig.width=8, fig.height=8,fig.cap="Accuracy and F1 score of the classifications made by our models compared with those made by widely used software", echo=FALSE, message=FALSE}
data.table::fread("data/processed/acc_f1score_vs_othertools.csv") %>%
  dplyr::mutate(model = case_when(
    model == "ensemble_weighted" ~ "Ensemble Weighted",
    model == "cnn_lstm" ~ "CNN-LSTM",
    model == "deepT3" ~ "DeepT3",
    model == "effectiveT3" ~ "EffectiveT3",
    model == "effectorp_1" ~ "EffectorP 1",
    model == "effectorp_2" ~ "EffectorP 2",
    TRUE ~ "EffectR"
  )) %>%
  dplyr::mutate(
    model = factor(model, levels = c("Ensemble Weighted", "DeepT3", "EffectiveT3",
                                    "CNN-LSTM", "EffectorP 1", "EffectorP 2",
                                    "EffectR"))
  ) %>%
  ggplot2::ggplot() +
  ggplot2::aes(
    x = model,
    y = .estimate,
    color = .metric,
    shape = .metric
  ) +
  ggplot2::scale_shape_manual(values = c("Accuracy" = 19, "F1 Score" = 17)) +
  ggplot2::geom_point(size = 2) +
  ggplot2::facet_wrap(~groups, nrow = 1, scales = "free_x") +
  ggplot2::scale_color_viridis_d(begin = 0.65, end = 0.1) +
  ggplot2::labs(x = "Classifier", y = "Score", color = "Metric", shape = "Metric") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

```{r vsotherscfmat, fig.width=8, fig.height=8, fig.cap="Confusion matrices of classifications made by our models compared with those made by widely used software. A: Tools evaluated on bacterial sequences; B: Tools evaluated on fungal sequences; C: Tools evaluated on oomycete sequences.", message = FALSE, echo=FALSE}

@@

```{r activations, fig.height=10, fig.width=10, fig.cap="Grad-CAM activation profiles of CNN-LSTM models trained on when run with all datasets; and sequence logo plots of sequence in the activation region within the dotted lines for effector data and non-effector data. A: Models trained on bacterial sequences; B: Models trained on fungal sequences; C: Models trained on oomycete sequences."}
# knitr::include_graphics("figures/convolution_maps_fixed.pdf")
